---
title: 'Lecture 3-3: The Spatial Dimension - Economic Complexity'
author: "Daniel S. Hain (dsh@business.aau.dk)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
    theme: flatly
---

```{r setup, include=FALSE}
### Generic preamble
Sys.setenv(LANG = "en") # System language = english
options(scipen = 5) # Disable scientific exponent annotation of numbers
set.seed(1337) # Set seed for reproducability

### Knitr options
library(knitr) # For display of the markdown
opts_chunk$set(warning=FALSE,
               message=FALSE,
               fig.align="center"
               )

### Load packages  
list.of.packages <- c("pacman")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages); rm(list.of.packages, new.packages)
```

```{r setup2}
### Load packages  
# Note: Maybe pacman needs to be installed
pacman::p_load(tidyverse # Collection of all the good stuff like dplyr, ggplot2 ect.
               , magrittr # For extra-piping operators (eg. %<>%)
               , tidygraph # For tidy-style graph manipulation
               , ggraph # For ggplot2 style graph plotting
               , ggmap # Geoplotting / coding
               )
```


# A closer look at the  Chinese patenting activity

## Loading and understanding the data structure
So, let's start the fun. I for you extracted Chinese patents from our EPO [PATSTAT](https://www.epo.org/searching-for-patents/business/patstat.html) databases filed at either the EPO or the USTPO. I further provide you adittional data. Lets take a look:

```{r}
patents <- readRDS(url("https://github.com/daniel-hain/SDC_IM/raw/master/data/CN_patent.rds"))
inventors <- readRDS(url("https://github.com/daniel-hain/SDC_IM/raw/master/data/CN_inventor.rds")) %>% filter(str_detect(city, "china")) %>% mutate(city = city %>% str_remove("\\,.*")) 
el_pat_inv <- readRDS(url("https://github.com/daniel-hain/SDC_IM/raw/master/data/CN_el_inventor_patent.rds"))
el_pat_tech <- readRDS(url("https://github.com/daniel-hain/SDC_IM/raw/master/data/CN_el_patent_field.rds"))
```

```{r}
field_names <- read_csv(url("https://github.com/daniel-hain/SDC_IM/raw/master/data/ipc_technology.csv")) %>%
  select(field_nr, sector, field_name) %>%
  distinct(field_nr, .keep_all = TRUE) %>%
  mutate(field_nr = field_nr) %>%
  arrange(field_nr)
```

```{r}
cities_geo <- inventors %>%
  distinct(city, .keep_all = TRUE) %>%
  select(city, lon, lat)
```

```{r}
reg_pat <- el_pat_inv %>%
  arrange(appln_id, invt_seq_nr) %>%
  distinct(appln_id, .keep_all = TRUE) %>%
  left_join(inventors %>% select(person_id, city), by = "person_id") %>%
  left_join(patents %>% select(appln_id, appln_filing_year, nb_citing_docdb_fam), by = "appln_id") %>%
  select(-invt_seq_nr, person_id) 
```

# Economic complexity of Chinese patenting activity

Let's not take a look at the complexity of Chinese patenting activity.

We will here use some neath functions from the`R` package `EconGeo`. Its only on github, so we have to install it from there.

```{r}
#remotes::install_github("PABalland/EconGeo", force = T)
library(EconGeo)
```

## Data munging
First, here we will need to create some big matrices, therefore I will here already define a little helper function which takes an edgelist as input and produces really efficient a sparse `M[i,j]` 2-mode matrix.

```{r}
## Helper function
create_sparse_matrix <- function(i.input, j.input){
  require(Matrix)
  mat <- spMatrix(nrow = i.input %>% n_distinct(),
                  ncol = j.input %>% n_distinct(),
                  i = i.input %>% factor() %>% as.numeric(),
                  j = j.input %>% factor() %>% as.numeric(),
                  x = rep(1, i.input %>% length() ) )
  
  row.names(mat) <- i.input %>% factor() %>% levels()
  colnames(mat) <- j.input %>% factor() %>% levels()
  return(mat)
}
```

Now, we will create this sparse matrix between patents and their associated technology classes.

```{r}
mat_tech <- create_sparse_matrix(i = el_pat_tech %>% pull(appln_id),
                                 j = el_pat_tech %>% pull(techn_field_nr))
```

Reminding basic matrix algebra, a 2-mode matrix `M[i,j]` can be transformed to a 1-mode adjacency matrix `M[j,j]` by taking the dot-product (matrix multiplication) of its transposed version `t(M[i,j])`, and into a  1-mode adjacency matrix `M[i,i]` by taking the dot-product with itself. In summary: `M[i,j] %*% M[i,j] = M[j,j]` and `M[i,j] %*% t(M[i,j]) = M[i,i]`. 

We can do that very efficiently with the `crossprod()` and `tcrossprod()` function of the `Matrix` package.

```{r}
mat_tech %<>% 
  crossprod() %>% 
  as.matrix() 
#mat.pat <- mat.tech %>% tcrossprod() %>% as.matrix() # Would create a patent matrix based on shared IPC classes (Not a good idea here)
```

```{r}
mat_tech
```


This matrix can be used for the `relatedness()` function to transform the co-occurence adjacency matrix to a similarity matrix. This function computes the relatedness (Hidalgo et al., 2007; Boschma et al., 2015; Balland, 2016) between entities (industries, technologies,...) from their co-occurrence (adjacency) matrix. Different normalization procedures are proposed following van Eck and Waltman (2009): association strength, cosine, Jaccard, and an adapted version of the association strength (probability index). I here choose the cosine similarity, since it is not so sensitive to scale.

```{r}
mat_tech_rel <- mat_tech %>% 
  relatedness(method = "cosine")
```

Lets check what happened?

```{r}
mat_tech_rel[1:10,1:10]
```

Now, at the row-column intercept, we will not have the number of co-occurences, but the relatedness index instead.

## The Chinese Technology Space
With this matrix, we can now create the Chinese "Technology Space" a'la Hidalgo & Hausmann (2007). Here we create a network of technology fields. I will delete below average weights to create some sparsity, and calculate the eigenvector centrality. I will here use the Eigenvector centrality as an easy proxi for technological complexity.

```{r}
library(tidygraph)

g_tech <- mat_tech_rel %>% as_tbl_graph(directed = FALSE) %N>%
  left_join(field_names %>% mutate(field_nr = field_nr %>% as.character()), by = c("name" = "field_nr")) %>%
  mutate(dgr = centrality_eigen(weights = weight)) %E>%
  filter(weight >= mean(weight))
```

```{r}
g_tech 
```


We will straight away visualize it as a network. We scale the nodes by their centrality, and color them by their higher level tech-sector.

In order to be able to reproduce the same node position in comparative graphs lateron, we will first save the node layout (classical fruchterman-reingold layout).


```{r}
coords_tech <- g_tech %>% igraph::layout.fruchterman.reingold() %>% as_tibble()
colnames(coords_tech) <- c("x", "y")
```


```{r,fig.width=12.5,fig.height=7.5}
library(ggraph)
g_tech %>%
  ggraph(layout =  coords_tech) + 
  geom_edge_link(aes(width = weight), alpha = 0.2, colour = "grey") + 
  geom_node_point(aes(colour = sector, size = dgr)) + 
  geom_node_text(aes(label = field_name), size = 4, repel = TRUE) +
  theme_graph()
```

Lets see which tech-field is the most central (complex) one:

```{r}
g_tech %N>% 
  arrange(desc(dgr)) %>%
  as_tibble() %>%
  slice(1:10)
```

And the least one...

```{r}
g_tech %N>% 
  arrange(dgr) %>%
  as_tibble() %>%
  slice(1:10)
```


Optimally, one would have created this space on world-data, and not only Chinese patents, though. Let's, now that we know the overall space, take a look at city level specialization and complexity.

## Specialization pattern of cities
We here first join our city level patent data with the technology fields. Since a patent can have multiple ones, we this time fractionalize it with a `field_weight`.

```{r}
reg_tech <- reg_pat %>%
  left_join(el_pat_tech %>% select(appln_id, techn_field_nr), by = "appln_id") %>%
  group_by(appln_id) %>%
  mutate(field_weight = 1 / n()) %>%
  ungroup()
```

Now we first aggregate it on the whole time frame. We could for sure also do that sepperate for different years. We filtr our cities with 

```{r}
reg_tech %<>%
  group_by(city, techn_field_nr) %>%
  summarise(n_tech_reg = sum(field_weight)) %>%
  ungroup() %>%
  drop_na() 
```

Now, we for the `EconGeo` package again have to transform it to a matrix. Again, we have to set the rownames, which are depreciated in the `tibble`

```{r}
mat_reg_tech <- reg_tech %>%
  arrange(techn_field_nr, city) %>%
  pivot_wider(names_from = techn_field_nr, values_from = n_tech_reg, values_fill = list(n_tech_reg = 0))

rn_save <- mat_reg_tech %>% pull(city)

mat_reg_tech %<>% select(-city) %>%
  as.matrix() %>%
  round()

rownames(mat_reg_tech) <- rn_save; rm(rn_save)
```

Lets take a look:

```{r}
mat_reg_tech[1:10, 1:10]
```


Now, we can use the `location.quotient()` function from `EconGeo`. This function computes location quotients from (incidence) regions - industries matrices. The numerator is the share of a given industry in a given region. The denominator is the share of a this industry in a larger
economy (overall country for instance). This index is also referred to as the index of Revealed Comparative Advantage (RCA) following Ballasa (1965), or the Hoover-Balassa index. To make it simple, we create a binary one, only indicating if the city has a RCA in the associated field or not. Afterwards, we transform it back to a tibble, set rownames again in a column, and gather it from a long to wide format.

```{r}
reg_RCA <- mat_reg_tech %>% location.quotient(binary = TRUE) %>% 
  as.data.frame() %>% 
  rownames_to_column("city") %>% 
  as_tibble() %>% 
  gather(key = "tech_field", value = "RCA", -city) %>%
  arrange(city, tech_field)
```

```{r}
reg_RCA %>% head()
```

The `Herfindahl()` function computes the Herfindahl index from regions - industries matrices from (incidence) regions - industries matrices. It is a measure of concentration. This index is also known as the Herfindahl-Hirschman index (Herfindahl, 1959; Hirschman, 1945).

```{r}
mat_reg_tech %>% 
  Herfindahl() %>% 
  as.data.frame() %>% 
  rownames_to_column("city") %>% 
  rename(HH = ".") %>% 
  arrange(desc(HH)) %>% 
  head(10)
```

Next, we can compute the Shannon entropy index (Shannon and Weaver, 1949; Frenken et al., 2007) from regions - industries matrices from (incidence) regions - industries matrices.

```{r}
mat_reg_tech %>% 
  entropy() %>% 
  as.data.frame() %>% 
  rownames_to_column("city") %>% 
  rename(SE = ".") %>% 
  arrange(desc(SE)) %>%
  head(10)
```

Now, finally, we can inspect where in the technology space different cities are specialized. We select some of interest.

```{r}
city_select <- c("beijing", "shanghai", "shenzhen")
```

In the following, I  plot the technology-space and highlights in red the fields where the city has an RCA. 

### Beijing

```{r,fig.width=12.5,fig.height=7.5}
i = 1
g_tech %N>%
  left_join(reg_RCA %>% filter(city == city_select[i]) %>% select(-city), by = c("name" = "tech_field")) %>%
  ggraph(layout = coords_tech) + 
  geom_edge_link(aes(width = weight), alpha = 0.2, colour = "grey") + 
  geom_node_point(aes(colour = RCA, size = dgr)) + 
  geom_node_text(aes(label = field_name), size = 5, repel = TRUE) +
  scale_color_gradient(low = "skyblue", high = "red") +
  theme_graph() +
  ggtitle(paste("Technology Space: RCA", city_select[i], sep = " "))
```


### Shanghai

```{r,fig.width=12.5,fig.height=7.5}
i = 2
g_tech %N>%
  left_join(reg_RCA %>% filter(city == city_select[i]) %>% select(-city), by = c("name" = "tech_field")) %>%
  ggraph(layout = coords_tech) + 
  geom_edge_link(aes(width = weight), alpha = 0.2, colour = "grey") + 
  geom_node_point(aes(colour = RCA, size = dgr)) + 
  geom_node_text(aes(label = field_name), size = 5, repel = TRUE) +
  scale_color_gradient(low = "skyblue", high = "red") +
  theme_graph() +
  ggtitle(paste("Technology Space: RCA", city_select[i], sep = " "))
```


### Shenzhen

```{r,fig.width=12.5,fig.height=7.5}
i = 3
g_tech %N>%
  left_join(reg_RCA %>% filter(city == city_select[i]) %>% select(-city), by = c("name" = "tech_field")) %>%
  ggraph(layout = coords_tech) + 
  geom_edge_link(aes(width = weight), alpha = 0.2, colour = "grey") + 
  geom_node_point(aes(colour = RCA, size = dgr)) + 
  geom_node_text(aes(label = field_name), size = 5, repel = TRUE) +
  scale_color_gradient(low = "skyblue", high = "red") +
  theme_graph() +
  ggtitle(paste("Technology Space: RCA", city_select[i], sep = " "))
```

So, what differences do we see?

# Endnotes

### References

### More info
You can find more info about:

#### Data
* [`Economic Complexity Data`](https://atlas.cid.harvard.edu/)

### Session info
```{r}
sessionInfo()
```



